Open Data for _Crime and Place_ Research: A Practical Guide in R.

### 2020-11-16 draft_chapter.Rmd
update stratford example & pre print version

## Cleaning, wrangling, and visualising our data

@@

```{r mapline}
ggplot() +
  geom_sf(data = tfl_jub_sf) +
  geom_sf(data = osm_jub_sf, color = "red", alpha = 0.5) 
```

As we can see, the two data sources appear to be broadly comparable in terms of their spatial patterning. However, the exact location of stations differs between Open Street Map and TfL. Determining a single pinpoint location for underground stations is not straightforward, especially for larger stations which have multiple entrances, platforms, and walkway tunnels. For operational purposes, TfL might have a specific definition for the location of a station, such as its centroid, which has been systematically applied to create the coordinates. The data pulled from Open Street Map might have been compiled by a volunteer contributor with their own perception of how to define the location of a station, such as the ticket booth or the main entrance. This definition _might_ be more theoretically meaningful than that used by TfL, since it could reflect a realistic 'on the ground' perception of the station's built environment, and thus opportunities for crime. On the other hand, if the locations from Open Street Map were collated by multiple different contributors, each with their own idea of where each station lies on the line, we might conclude that the measure is unreliable. These kinds of questions demonstrate why critically engaging with open data sources is so crucial, but also challenging. We return to this at the end of the exercise.

@@

# Visualize the building footprints, Open Street Map buffer, TfL buffer, and crime
# locations. Note that the single point for the crimes is due to the approximation!
# There are actually multiple crimes.
ggplot() +
  theme_bw() +
  geom_sf(data = strat_fp_poly_sf, fill = "black") +
  geom_sf(data = stratford_osm_sf, fill = "pink", alpha = 0.3) +
  geom_sf(data = stratford_tfl_sf, fill = "blue", alpha = 0.3) +
  geom_sf(data = strat_osm_btp_sf, col = "red") 
```

@@

This quick check has shed some light on what is happening. The buffer generated from the TfL station location (in blue) covers a different part of the station to the Open Street Map buffer (in pink). The latter covers most of the main station building, while the former centers around the platforms to the north. Because the British Transport Police crime locations have been geocoded to a location in the main station building (in red), the TfL buffer has missed them. If we had used a larger buffer, it is likely that the TfL location would have captured the crime cluster, but then we would increase the risk of picking up crimes outside of the station. In many cases, there will not be an obvious 'right or wrong' answer to determine which open data source is most appropriate, but investigations like this one will help inform the decision. We might consider investigating other stations before deciding on which data to use, and ensure that findings are reported transparently with reference to these discussions.

@@

>>>>

## Cleaning, wrangling, and visualizing our data

@@

```{r mapline}
ggplot() +
  geom_sf(data = tfl_jub_sf) +
  geom_sf(data = osm_jub_sf, col = "red", alpha = 0.5) 
```

As we can see, the two data sources appear to be broadly comparable in terms of their spatial patterning. However, the exact location of stations sometimes differs between Open Street Map and TfL. Determining a single pinpoint location for underground stations is not straightforward, especially for larger stations which have multiple entrances, platforms, and walkway tunnels. For operational purposes, TfL might have a specific definition for the location of a station, such as its centroid, which has been systematically applied to create the coordinates. The data pulled from Open Street Map might have been compiled by a volunteer contributor with their own perception of how to define the location of a station, such as the ticket booth or the main entrance. This definition _might_ be more theoretically meaningful than that used by TfL, since it could reflect a realistic 'on the ground' perception of the station's built environment, and thus opportunities for crime. On the other hand, if the locations from Open Street Map were collated by multiple different contributors, each with their own idea of where each station lies on the line, we might conclude that the measure is unreliable. These kinds of questions demonstrate why critically engaging with open data sources is so crucial, but also challenging. We return to this at the end of the exercise.

@@

# Visualize the building footprints, Open Street Map buffer, TfL buffer, and crime
# locations. Note that the single point for the crimes is due to the approximation!
# There are actually multiple crimes. Colors have been manually selected
# from the viridis package to be easier to read by those with colorblindness.
ggplot() +
  geom_sf(data = strat_fp_poly_sf, fill = "black") +
  geom_sf(data = stratford_osm_sf, fill = "#440154FF", alpha = 0.3) +
  geom_sf(data = stratford_tfl_sf, fill = "#FDE725FF", alpha = 0.3) +
  geom_sf(data = strat_osm_btp_sf, col = "#3CBC75FF", fill = "white") +
  theme(axis.text = element_text(size = 6))
```

@@

This quick check has shed some light on what is happening. The buffer generated from the TfL station location (in yellow) covers a different part of the station to the Open Street Map buffer (in blue). The latter covers most of the main station building, while the former centers around the platforms to the north. Because the British Transport Police crime locations have been geocoded to a location in the main station building (in green) the TfL buffer has missed them. If we had used a larger buffer, it is likely that the TfL location would have captured the crime cluster, but then we would increase the risk of picking up crimes outside of the station. In many cases, there will not be an obvious 'right or wrong' answer to determine which open data source is most appropriate, but investigations like this one will help inform the decision. We might consider investigating other stations before deciding on which data to use, and ensure that findings are reported transparently with reference to these discussions.

@@

# Software

As noted in the tutorial, a number of R packages were used to create the material in this chapter. Packages _readr_ [@readrpack] and _jsonlite_ [@jsonlitepack] were used for reading and parsing data, _osmdata_ for querying the Open Street Map API [@osmdatapack], _sf_ for handling and visualizing spatial data [@sfpack], _dplyr_ and _tidyr_ for wrangling data [@dplyrpack; @tidyrpack], _ggplot2_ for visualization [@ggplotpack] and _patchwork_ for arranging plots [@patchworkpack]. The document itself was prepared using _rmarkdown_ [@rmarkdownpack].
